### 2021-08-13 docs/revision/appendix_b.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\hypertarget{missing-data}{%
\section{Missing data}\label{missing-data}}

Of the 78 experiments that met the eligibility criteria of this
meta-analysis, 25 were excluded because of missing data. Those 25
experiments included 13 experiments that reported a statistically
significant result, along with 12 that failed to find a significant
self-controlled learning effect. Among the 13 experiments with missing
data reporting a significant self-control benefit, one reported an
inappropriate analysis (Hemayattalab et al.2013),\footnote{Although data
  were collected in one dimension using concentric circles, AE and a
  measure of dispersion were analyzed together in a MANOVA. This measure
  of dispersion is not an accurate reflection of variability on a
  two-dimensional task for reasons described by Hancock et al.~(1995).}
one reported statistics that do not match the experimental design
(Jalalvand et al., 2019),\footnote{A subgroup analysis involving two
  groups \(n\) = 15 was reported with \(df\) = 56. The article reports
  \(r^2\) effect sizes associated with each test that cannot be
  reproduced with the reported statistics or best guesses.} one reported
significant effects on a partial analysis of their data rather than
overall (Brydges et al., 2009), and one was previously identified by
Lohse and colleagues (2016) as an outlier study (Carter \& Patterson,
2012). The meta-analysis may have been strengthened by the exclusion of
these results (Stanley et al., 2010).

Among the remaining nine experiments reporting a significant effect with
missing data, two reported effects collapsed across immediate and
delayed retention only (Patterson et al., 2013; Wu \& Magill, 2011), two
reported null effects on a higher priority measure and did not include
sufficient data to calculate the effect size, while reporting a
significant effect on a lower priority measur((Wulf et al., 2001; Wulf
et al., 2005, both studies were included in the primary \emph{p}-curve
analysis), and five compared three or more groups in an omnibus ANOVA
and reported the group effect as significant but did not include
sufficient data to calculate the effect size for the self-control versus
yoked comparison (Chen et al., 2002; Ghorbani, 2019; Huet et al., 2009;
Janelle et al., 1997; Norouzi et al., 2016)
\end{appendix}